---
title: "Why Don't Americans Vote"
author: "Oliver Schaack, Dylan Jaffe, Michael Stamatos"
date: "October 6th, 2024"
output:
  html_document:
    theme: cerulean
    highlight: pygments
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction 

In this notebook, we are reverse engineering the story, [Why Don't Americans Vote](https://projects.fivethirtyeight.com/non-voters-poll-2020-election/)

## Load libraries

Loading required libraries for this analysis.

```{r echo=FALSE, message=FALSE}
library(tidyverse)
library(lubridate)
library(janitor)
library(refinr)
```

## Load and Cleaning Data

In this section, describe the source of the data, write a basic data dictionary for data you are working with, and discuss any caveats or issues you discovered working with this data. 

```{r}
# Load required data

# Path to data should be loaded from folder "data" i.e. read_csv("data/name_of_data.csv")
Story_data <- read_csv("data/nonvoters_data.csv")
view(Story_data)
# Clean required data and prepare for analysis if needed. 


```

## Sentences to Engineer
In this notebook, we are reverse engineering five sentences from the story.

### Sentence 1

* **Sentence text**: [Every election, millions of Americans go through a similar thought process and, it turns out, lots of people feel like Brown: They think voting doesn’t matter or isn’t worth their time.]
* **Analysis summary**: [611 people fit that category out of an original dataset of 5,836 - roughly 10 percent, which would equate to millions of people with the correlation to the US voting population]

```{r}
# Put code to reverse engineer sentence here
filtered_data<-Story_data |>
filter(Q26 == 2 | Q26 == 3 | Q26 == 4)
Story_data
# Display results of code below this codeblock

```
```{r}
filtered_data <- filtered_data |>
  filter(Q29_2 == 1 | Q29_3 == 1 | Q29_4 == 1)
filtered_data
```

### Sentence 2

* **Sentence text**: [In the survey, we asked voters who have missed at least one national election — which included some people who almost always vote — why they didn’t cast a ballot. Nearly a quarter cited some of the structural barriers we mentioned above. But another 31 percent said that they decided not to vote because they disliked the candidates or they thought nothing would change as a result of the election (26 percent)]
* **Analysis summary**: [Among the 1,342 people who say they sometimes, rarely, or don't vote in national elections, 765 say they either dislike the candidates or that nothing will change for them. 403 say they dislike the candidates, while 362 say nothing will change as a result, which is similar to the original numbers presented, where 403 is 30% of 1342 and 362 is 27%. ]

```{r}
# Put code to reverse engineer sentence here
filtered_data <- Story_data |>
  filter(Q26 == 2 | Q26 == 3 | Q26 == 4)

filtered_data
# Display results of code below this codeblock
filtered_data <- filtered_data |>
  filter(Q29_1 == 1 | Q29_3 == 1)
print(filtered_data)
```

### Sentence 3

* **Sentence text**: [“Williams is certainly not alone in experiencing these barriers. Of the three groups of voters we identified, those voters who only vote some of the time were actually the likeliest to report having stood in line for more than an hour; they were also likelier than those who vote more regularly to say they couldn’t get off work to vote.”]
* **Analysis summary**: 

```{r}
filtered_data <- Story_data |>
  filter(Q14 == 4 | Q14 == 5)

filtered_data <- filtered_data |>
  filter(Q29_1 | Q29_3 == 1)

filtered_data
```

### Sentence 4

* **Sentence text**: ["A substantial chunk of respondents said they don’t think either party wants people like them to vote. (Twenty-three percent said this of Democrats and 31 percent of Republicans.) Black and Hispanic people were especially likely to say the Republicans don’t want people like them to vote: Fifty-four percent of Black respondents and 35 percent of Hispanic respondents agreed with this statement, compared to just 26 percent of white respondents."]
* **Analysis summary**: [The results of the analysis confirm the findings pretty well. Q14, represeting the republican party, shows that 33% of correspondants don't think republicans want them to vote, which is similar to the 31% in the article. Q15 represents the Democratic party and got 21% but accounting for people that didn't vote the number should be similar to the 23% in the article.

```{r}
Story_data |>
  group_by(Q14) |>
  summarise(count = n()) |>
  mutate(percent = (count/sum(count)*100))
Story_data
```

### Sentence 5

* **Sentence text**: ["There are, of course, other systemic reasons why some people might vote more inconsistently. Our survey found, for instance, that occasional voters were slightly more likely than frequent voters to have a long-term disability (8 percent vs. 5 percent), and nonvoters were even more likely to fall into this category (12 percent). Black and Hispanic voters are also more likely to experience hurdles, perhaps in part because there tend to be fewer polling places in their neighborhoods. About 24 percent of Black respondents said that they had to stand in line for more than an hour while voting, and Hispanic respondents were more likely to say they had trouble accessing the polling place or couldn’t get off work in time to vote.”]
* **Analysis summary**: [248/392, or 26% of black respondants, said that they had to stand in line for more than an hour while voting, which is similar to the 24% number in the article. In addition, 130/813 Hispanics said they had trouble accessing the polling place or could not get off work in time to vote.]

```{r}
filtered_data <- Story_data
filtered_data |>
  filter(race == "Black") |>
  filter(Q18_8 == 1) 
print(filtered_data)

filtered_data <- Story_data
filtered_data |>
  filter(race == 'hispanic') |>
  filter(Q18_2 == 1 | Q18_7 == 1)
print(filtered_data)
```

-30-