---
title: "Why Don't Americans Vote"
author: "Oliver Schaack, Dylan Jaffe, Michael Stamatos"
date: "October 6th, 2024"
output:
  html_document:
    theme: cerulean
    highlight: pygments
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction 

In this notebook, we are reverse engineering the story, [Why Don't Americans Vote](https://projects.fivethirtyeight.com/non-voters-poll-2020-election/)

## Load libraries

Loading required libraries for this analysis.

```{r echo=FALSE, message=FALSE}
library(tidyverse)
library(lubridate)
library(janitor)
library(refinr)
```

## Load and Cleaning Data

In this section, describe the source of the data, write a basic data dictionary for data you are working with, and discuss any caveats or issues you discovered working with this data. 

```{r}
# Load required data

# Path to data should be loaded from folder "data" i.e. read_csv("data/name_of_data.csv")
Story_data <- read_csv("data/nonvoters_data.csv")
view(Story_data)
# Clean required data and prepare for analysis if needed. 


```

## Sentences to Engineer
In this notebook, we are reverse engineering five sentences from the story.

### Sentence 1

* **Sentence text**: [Every election, millions of Americans go through a similar thought process and, it turns out, lots of people feel like Brown: They think voting doesn’t matter or isn’t worth their time.]
* **Analysis summary**: [611 people fit that category out of an original dataset of 5,836 - roughly 10 percent, which would equate to millions of people with the correlation to the US voting population]

```{r}
# Put code to reverse engineer sentence here
filtered_data<-Story_data |>
filter(Q26 == 2 | Q26 == 3 | Q26 == 4)
Story_data
# Display results of code below this codeblock

```
```{r}
filtered_data <- filtered_data |>
  filter(Q29_2 == 1 | Q29_3 == 1 | Q29_4 == 1)
filtered_data
```

### Sentence 2

* **Sentence text**: [In the survey, we asked voters who have missed at least one national election — which included some people who almost always vote — why they didn’t cast a ballot. Nearly a quarter cited some of the structural barriers we mentioned above. But another 31 percent said that they decided not to vote because they disliked the candidates or they thought nothing would change as a result of the election (26 percent)]
* **Analysis summary**: ["When filtering the voter category, the largest amount of responses said they rarely/ never vote, showing that many people believe their vote does not matter. I filtered for questions that asked voters how often they vote and why they did not vote. Out of 1,342 people who say they sometimes, rarely, or don't vote in national elections, 765 say that they either dislike the candidates or that nothing will change for them. 403 say they dislike the candidates, while 362 say nothing will change as a result. This is similar to the original numbers presented in the text, as 403 is 30% of 1342 and 362 is"]

```{r}
# Put code to reverse engineer sentence here
grep("Q26", names(Story_data), value = TRUE)
```
```{r}
stwo_data <- Story_data |>
  filter(!is.na(Q26), !is.na(Q29_1), !is.na(Q29_2), !is.na(Q29_3)) |>
  group_by(voter_category) |>
  summarise(count=n())
stwo_data
```

### Sentence 3

* **Sentence text**: [“They were also likelier than those who vote more regularly to say they couldn't get off work to vote..Also because he doesn't trust mail-in ballots. "The last time I tried to vote by mail, it didn't get counted - it was too late," he said. As a result, Sanchez told us, "I don't know where my ballot is going to end up, if I don't see it put into the voting machine.".”]
* **Analysis summary**: Question 29 provided ten answer selections for study participants to choose from when explaining why they did not vote in the most recent presidential election, which, in this case, was the 2020 presidential election. Selection 5, also referred to as Q29_5 in our code, states, "I wanted to vote, but I didn't have time, couldn't get off work, something came up, or I forgot." Our sentence above for this analysis suggests that "they," referring to people who never vote, are more likely to abstain from voting compared to individuals who vote more frequently, labeled as the "sporadic group," because they could not get off work, which falls under selection 5 for Question 29. A response of -1 and N/A in the Q29_5 column indicates people who did not answer the question, while a response of 1 represents those who did, categorized as either rarely/never voters or sporadic voters. Approximately 10.13% of rarely/never voters selected option 5 for their reason for not voting, whereas only 2.72% of sporadic voters chose this option. This supports the article's assertion that around five times as many rarely/never voters selected option 5 compared to sporadic voters.


```{r}
Story_data |>
  filter(voter_category == "sporadic")|>
  group_by(Q29_5) |>
  summarise(count=n()) |>
  mutate(percentages = count/sum(count)*100)
  
```
```{r}
Story_data |>
  filter(voter_category == "rarely/never")|>
  group_by(Q29_5) |>
  summarise(count=n()) |>
  mutate(percentages = count/sum(count)*100)
```


### Sentence 4

* **Sentence text**: ["A substantial chunk of respondents said they don’t think either party wants people like them to vote. (Twenty-three percent said this of Democrats and 31 percent of Republicans.) Black and Hispanic people were especially likely to say the Republicans don’t want people like them to vote: Fifty-four percent of Black respondents and 35 percent of Hispanic respondents agreed with this statement, compared to just 26 percent of white respondents."]
* **Analysis summary**: The result of the analysis confirm the findings pretty well. 014, which represents the republican party, shows that 33% of respondents say they don't think republicans want them to vote, similar to the 31% in the article. 15, which is the democratic party, got a number of 21%, but accounting for the people that didn't vote the number should be similar to the 23% in the article. For the individual race numbers, I got 54% for black people and 29% for hispanics. 54 percent is right in line with
Files the 54 percent from the text, and the 29 is close enough to 34 considering a five percent margin of error.

```{r}
Hispanic_story_data <- Story_data |>
  filter(race == "Hispanic")

view(Hispanic_story_data)
```

```{r}
Hispanic_story_data |>
  group_by(Q14) |>
  summarise(count = n()) |>
  mutate(percent = (count/sum(count)*100))
Hispanic_story_data
```

```{r}
Black_story_data <- Story_data |>
  filter(race == "Black") 
view(Black_story_data)
```

```{r}
Black_story_data |>
  group_by(Q14) |>
  summarise(count = n()) |>
  mutate(percent = (count/sum(count)*100))
```

### Sentence 5

* **Sentence text**: ["There are, of course, other systemic reasons why some people might vote more inconsistently. Our survey found, for instance, that occasional voters were slightly more likely than frequent voters to have a long-term disability (8 percent vs. 5 percent), and nonvoters were even more likely to fall into this category (12 percent). Black and Hispanic voters are also more likely to experience hurdles, perhaps in part because there tend to be fewer polling places in their neighborhoods. About 24 percent of Black respondents said that they had to stand in line for more than an hour while voting, and Hispanic respondents were more likely to say they had trouble accessing the polling place or couldn’t get off work in time to vote.”]
* **Analysis summary**: [248/392, or 26% of black respondants, said that they had to stand in line for more than an hour while voting, which is similar to the 24% number in the article. In addition, 130/813 Hispanics said they had trouble accessing the polling place or could not get off work in time to vote.]

```{r}
filtered_data <- Story_data
filtered_data |>
  filter(race == "Black") |>
  filter(Q18_8 == 1) 
print(filtered_data)
```

```{r}
filtered_data <- Story_data
filtered_data |>
  filter(race == 'hispanic') |>
  filter(Q18_2 == 1 | Q18_7 == 1)
print(filtered_data)
```


-30-